---
title: "Lab4"
author: "Adam Brandt"
date: "10/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readxl)
library(tidyverse)
defense <- read_excel('./cyclonesFootball2019.xlsx', sheet='Defensive')
offense <- read_excel('./cyclonesFootball2019.xlsx', sheet='Offensive')
biography <- read_excel('./cyclonesFootball2019.xlsx', sheet='Biography')
```

### Part One: Cleaning data ###

1.
```{r}
player_names <- factor(biography$Name)
opponent_names <- factor(unique(defense$Opponent_Opponent))
```

2.
```{r}
offClean <- offense %>% separate(`Passing_CMP-ATT`, into = c("cmp", "att"), sep = '-')
offClean$cmp <- str_replace(offClean$cmp, ' ', '')
offClean$att <- str_replace(offClean$att, ' ', '')
offClean <- offClean %>% mutate(Receiving_REC = as.numeric(Receiving_REC))
offClean <- offClean %>% mutate(Receiving_YDS = as.numeric(Receiving_YDS))
offClean <- offClean %>% mutate(Receiving_TD = as.numeric(Receiving_TD))
offClean <- offClean %>% mutate(Rushing_ATT = as.numeric(Rushing_ATT))
offClean <- offClean %>% mutate(Rushing_YDS = as.numeric(Rushing_YDS))
offClean <- offClean %>% mutate(Rushing_TD = as.numeric(Rushing_TD))
offClean <- offClean %>% mutate(cmp = as.numeric(cmp))
offClean <- offClean %>% mutate(att = as.numeric(att))
offClean <- offClean %>% mutate(Passing_YDS = as.numeric(Passing_YDS))
offClean <- offClean %>% mutate(Passing_TD = as.numeric(Passing_TD))
offClean <- offClean %>% mutate(Passing_INT = as.numeric(Passing_INT))


defClean <- defense
defClean <- defClean %>% mutate(Tackles_Solo, Tackles_Solo = as.numeric(Tackles_Solo))
defClean <- defClean %>% mutate(Tackles_ASST, Tackles_ASST = as.numeric(Tackles_ASST))
defClean <- defClean %>% mutate(Tackles_TFL, Tackles_TFL = as.numeric(Tackles_TFL))
defClean <- defClean %>% mutate(Tackles_Sack, Tackles_Sack = as.numeric(Tackles_Sack))
defClean <- defClean %>% mutate(Turnover_FF, Turnover_FF = as.numeric(Turnover_FF))
defClean <- defClean %>% mutate(Turnover_FR, Turnover_FR = as.numeric(Turnover_FR))
defClean <- defClean %>% mutate(Turnover_INT, Turnover_INT = as.numeric(Turnover_INT))
defClean <- defClean %>% mutate(Pass_QBH, Pass_QBH = as.numeric(Pass_QBH))
defClean <- defClean %>% mutate(Pass_PB, Pass_PB = as.numeric(Pass_PB))
offClean
defClean
```

3.
```{r}
bioClean <- biography
bioClean <- bioClean %>% mutate(Weight, Weight = as.numeric(Weight))
bioClean <- bioClean %>% separate(Height, into = c('ft', 'inch'), sep = '-') %>%
  mutate(Height = as.numeric(ft)*12 + as.numeric(inch))
bioClean <- subset(bioClean, select = -c(ft, inch))
bioClean
```

### Part Two: Tidying ###

1.
```{r}
defClean <- defClean %>% pivot_longer(c('Tackles_Solo',	'Tackles_ASST',	'Tackles_TFL',	'Tackles_Sack',	'Turnover_FF',	'Turnover_FR',	'Turnover_INT',	'Pass_QBH',	'Pass_PB'),  names_to = "stat", values_to = "Recorded")
defClean
```

2.
```{r}
defClean %>% ggplot(aes(x=Recorded)) + geom_histogram() + facet_wrap(~stat)
```

The rarest stats are turnover interception, turnover fumble recovery and turnover forced fumble.

3.
```{r}
Iowa <- defClean %>% filter(Opponent_Opponent == 'Iowa', stat == 'Tackles_Solo')
Notre_Dame <- defClean %>% filter(Opponent_Opponent == 'Notre Dame', stat == 'Tackles_Solo')
tackles <- inner_join(Iowa, Notre_Dame, by='Name')
summary(tackles)
ggplot(tackles, aes(x=`Recorded.x`, y = `Recorded.y`)) + geom_point() + xlab('Iowa') + ylab('Notre Dame') + ggtitle("Solo Tackles against Iowa vs. Solo Tackles against Notre Dame")
```

The defense against Notre Dame was better since it had a higher average number of tackles.

4.
```{r}
bioClean <- bioClean %>% separate(Hometown, into = c('city', 'state'), sep = ', ') 
head(bioClean)
```

5.
```{r}
table(bioClean$state)
```


### Part Three: Joining Data Frames ###
1.
```{r}
joinOff <- inner_join(bioClean, offClean, by = "Name")
table(joinOff$state)
joinDef <- inner_join(bioClean, defClean, by = "Name")
table(joinDef$state)
```

2.
```{r}
Purdy <- offClean %>% filter(Name == "Purdy, Brock")
summary(Purdy)
summary(offClean)
```

Purdy performs above average compared to the rest of the players. For example he has a mean if 2.077 passing touchdowns where as the team as a mean of 0.935 passing touchdowns.

3.
```{r}
defense2 <- read_excel('./cyclonesFootball2018.xlsx', sheet='Defensive')
offense2 <- read_excel('./cyclonesFootball2018.xlsx', sheet='Offensive')
defClean2 <- defense2 %>% mutate(Tackles_Solo = as.numeric(Tackles_Solo))

offClean2 <- offense2 %>% mutate(Receiving_YDS = as.numeric(Receiving_YDS))
offComp <- inner_join(offClean, offClean2, by='Name') %>% 
           mutate(Receiving_YDS_Diff = Receiving_YDS.x - Receiving_YDS.y) %>% 
           select(Name, Receiving_YDS_Diff) %>% 
           arrange(-Receiving_YDS_Diff)
offComp
defClean2 <- defense %>% mutate(Tackles_Solo = as.numeric(Tackles_Solo))
defClean3 <- defense2 %>% mutate(Tackles_Solo = as.numeric(Tackles_Solo))
defComp <- inner_join(defClean2, defClean3, by='Name') %>% 
           mutate(Tackles_Solo_Diff = Tackles_Solo.x - Tackles_Solo.y) %>% 
           select(Name, Tackles_Solo_Diff) %>% 
           arrange(-Tackles_Solo_Diff)
defComp

```
On the defense Braxton Lewis and Anthony Johnson improved a lot and on the offense Deshaunte Jones and Tarique Milton improved a lot
